# Data

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

For our project, a prediction of the Premier League rankings, we have several datasets.

The first one is called `game_results` containing the results of 4'560 Premier League games (380 games over 12 seasons from 2006/2007 to 2017/2018):


```{r echo=FALSE, include=FALSE}
# Importing the games_results dataset into R
setwd(here::here(("Data/")))

game_results <- readr::read_csv("results.csv")
```

```{r, echo=FALSE}
# Show the dataset

game_results[1:5,] %>% 
  rename("home team" = home_team,
         "away team" = away_team,
         "home goals" = home_goals,
         "away goals" = away_goals) %>% 
  knitr::kable(caption = "**Table 1: Premier League results (from 2006 to 2018)**") %>% 
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                fixed_thead = T)
```

The second dataset `team_stats` is about the statistics of each team in every season from 2006/2007 to 2017/2018. The statistics can be split into 5 categories: general, attack, defence, team play and others.

```{r, echo=FALSE, include=FALSE}
# Importing the team_stats dataset into R
setwd(here::here(("Data/")))

team_stats <- readr::read_csv("stats.csv")
```

```{r, echo=FALSE}
# Show the dataset

team_stats[1:5,] %>% 
  select(team, wins, losses, goals, total_yel_card, total_scoring_att, saves, season) %>% 
  rename("total yellow cards" = total_yel_card,
         "total scoring attempts" = total_scoring_att) %>% 
  knitr::kable(caption = "**Table 2: Premier League teams statistics (from 2006 to 2018)**") %>% 
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                fixed_thead = T)
  
```

The third dataset `players_stats` is about the statistics of the players of Premier League until september 2020.

```{r, echo=FALSE, include=FALSE}
# Importing the players_stats dataset into R
setwd(here::here(("Data/")))

players_stats <- readr::read_csv("dataset - 2020-09-24.csv")
```

```{r, echo=FALSE}
players_stats[1:5,] %>% 
  select(Name, Club, Position, Appearances, Wins, Passes, Saves) %>% 
  knitr::kable(caption = "**Table 3: Premier League players statistics (until 2020)**") %>% 
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                fixed_thead = T)
  
```

The last dataset `players_value` contains information about the market value for the season 2016/2017

```{r, echo=FALSE, include=FALSE}
# Importing the players_value dataset into R
setwd(here::here(("Data/")))

players_value <- readr::read_csv("england_premier_league_players_data.csv")
```

```{r, echo=FALSE}

```

