# Data

```{r, echo = FALSE, include = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```
<div>
For our project, we will be providing a prediction of the Premier League rankings. We will be using several datasets to which we will apply different models to get a prediction then test its accuracy.

The first dataset is called `game_results` containing the results of 4'560 Premier League games (380 games over 12 seasons from 2006/2007 to 2017/2018). The below are the variables in the dataset and their details:
</div>
 + `home_team` - Team playing in their home stadium.
 + `away_team` - Team playing in the stadium of their opponent.
 + `home_goals` - Number of goals scored by the home team.
 + `away_goals` - Number of goals scored by the away team
 + `results` - Winner of the match (H-home team, A-away team, D-draw)
 + `season` - Mentions the season in which the match took place


```{r echo=FALSE, include=FALSE}
# Importing the games_results dataset into R
setwd(here::here(("Data/")))

game_results <- readr::read_csv("results.csv")
```

```{r, echo=FALSE}
# Show the dataset

game_results[1:5,] %>% 
  rename("home team" = home_team,
         "away team" = away_team,
         "home goals" = home_goals,
         "away goals" = away_goals) %>% 
  knitr::kable(caption = "**Table 1: Premier League results (from 2006 to 2018)**") %>% 
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                fixed_thead = T)
```

```{r}
# Write CSV files

setwd(here::here("data/"))
write_csv(game_results,"game_results_cleaned.csv")
```