# Data

```{r, echo = FALSE, include = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```
<div>
For our project, we will be providing a prediction of the Premier League rankings. We will be using several datasets to which we will apply different models to get a prediction then test its accuracy.

The dataset is called xxx containing the results of Premier League games over (380 games over 12 seasons from 2006/2007 to 2017/2018). The below are the variables in the dataset and their details:
</div>
 + `xxx` - ...
 + `xxx` - ...
 + `xxx` - ...
 + `xxx` - ...
 + `xxx` - ...
 + `xxx` - ...


```{r echo=FALSE, include=FALSE}
# Importing the games_results dataset into R
setwd(here::here(("Data_New/")))

premier_league_results_1718 <- readr::read_csv("season-1718_csv.csv")
premier_league_results_1819 <- readr::read_csv("season-1819_csv.csv")
premier_league_results_1718 <- subset(premier_league_results_1718, select = -c(LBH,LBD,LBA) )
premier_league_results <- full_join(premier_league_results_1718, premier_league_results_1819, by = c("Div","Date","HomeTeam","AwayTeam","FTHG","FTAG","FTR","HTHG","HTAG","HTR","Referee","HS","AS","HST","AST","HF","AF","HC","AC","HY","AY","HR","AR","B365H","B365D","B365A","BWH","BWD","BWA","IWH","IWD","IWA","PSH","PSD","PSA","WHH","WHD","WHA","VCH","VCD","VCA","Bb1X2","BbMxH","BbAvH","BbMxD","BbAvD","BbMxA","BbAvA","BbOU","BbMx>2.5","BbAv>2.5","BbMx<2.5","BbAv<2.5","BbAH","BbAHh","BbMxAHH","BbAvAHH","BbMxAHA","BbAvAHA","PSCH","PSCD","PSCA"))
```

```{r, echo=FALSE}
# Show the dataset

premier_league_results_1718[1:5,] %>% 
  rename("home team" = home_team,
         "away team" = away_team,
         "home goals" = home_goals,
         "away goals" = away_goals) %>% 
  knitr::kable(caption = "**Table 1: Premier League results (seasons 2017-18 and 2018-2019)**") %>% 
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                fixed_thead = T)
```

```{r}
# Write CSV files

setwd(here::here("data/"))
write_csv(xxx,"xxx")
```