<style>
body {
text-align: justify}
</style>

# Data

```{r, echo = FALSE, include = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

## Data sources

For this report, we will be using datasets representing the details of the Premier League games for three consecutive seasons; 2016-17, 2017-18, and 2018-19. Our reason for choosing three seasons is that we feel that the data for three years contains a sufficient number of observations without risking the change of non-quantitative data that is not captured in the datasets such as the players playing in the games and their skills. The source of our data was directly obtained from the data-hub website. The following are the links to each of the three datasets used:

1. https://datahub.io/sports-data/english-premier-league#resource-season-1617
2. https://datahub.io/sports-data/english-premier-league#resource-season-1718
3. https://datahub.io/sports-data/english-premier-league#resource-season-1819


## Data description 

The Premier League, also known as the English Premier League, is the top level of the English football league system. Contested by 20 clubs, it operates on a system of promotion and relegation with the English Football League. Seasons typically run from August to May with each team playing 38 matches. Each of the three datasets contains 380 observations, and each of them represents one match. The dataset for the season 2018-19 contains 62 variables, while the other two datasets contain 65 variables each. The variables are the following:

- general information like league division, match date and referee
- **home team**: team playing at their "own" field, receiving team
- **away team**: team playing at another field than their "own", visiting team
- **full time home goals**: the number of goals at the end of the match for the home team
- **full time away goals**: the number of goals at the end of the match for the away team
- **full time results**: match result at the end (H=Home win, A=away win, D=draw)
- **half time home goals**: the number of goals scored in the first part (45mins) of the game for the home team
- **half time away goals**: the number of goals scored in the first part (45mins) of the game for the away team
- **half time results**: match result for the first half of the game (H=Home win, A=away win, D=draw)
- **home shots**: ball hits as an attempt to score a goal by the home team
- **away shots**: ball hits as an attempt to score a goal by the away team
- **home shots ontarget**: home team shots that enters the goal or would have entered the goal if they had not been blocked by the goalkeeper or another defensive player
- **away shots ontarget**: away team shots that enters the goal or would have entered the goal if they had not been blocked by the goalkeeper or another defensive player, 
- **home fouls**: inappropriate or unfair acts by a player of the home team as deemed by a referee, that often result in a penalty
- **away fouls**: inappropriate or unfair acts by a player of the away team as deemed by a referee, that often result in a penalty
- **home corners**: awarded when the whole of the ball passes over the goal line, on the ground or in the air, having last touched a player of the defending team, and a goal is not scored
- **away corners**: awarded when the whole of the ball passes over the goal line, on the ground or in the air, having last touched a player of the defending team, and a goal is not scored
- **home yellow cards**: given to the home team as a caution or warning, provide players with another chance to stay on the field for the remainder of the game
- **away yellow cards**: given to the away team as a caution or warning, provide players with another chance to stay on the field for the remainder of the game
- **home red cards**: used by the officials to remove a home team player from the match with immediate effect
- **away red cards**: used by the officials to remove an away team player from the match with immediate effect
- and others like odds coming from various betting websites.


## Data cleaning

Before applying machine learning models to our dataset, we need to clean and transform it. We start by importing our data, taking out unknown variables, transforming non numerical variables to their correct type and joining the datasets from different seasons to ensure we have enough observations.

```{r echo=FALSE, include=FALSE}
# Importing the games_results dataset into R
setwd(here::here(("Data_New/")))

premier_league_results_1718 <- readr::read_csv("season-1718_csv.csv")
premier_league_results_1819 <- readr::read_csv("season-1819.csv")
premier_league_results_1617 <- readr::read_csv("season-1617.csv")

# Code the match_date
premier_league_results_1718 <- premier_league_results_1718 %>% 
  mutate(Date = as_date(Date, format="%d/%m/%y"))

premier_league_results_1819 <- premier_league_results_1819 %>% 
  mutate(Date = as_date(Date, format="%d/%m/%Y"))

premier_league_results_1617 <- premier_league_results_1617 %>% 
  mutate(Date = as_date(Date, format="%d/%m/%y"))

premier_league_results_1718 <- subset(premier_league_results_1718, select = -c(LBH,LBD,LBA,PSCH,PSCD,PSCA))
premier_league_results_1819 <- subset(premier_league_results_1819, select = -c(PSCH,PSCD,PSCA))
premier_league_results_1617 <- subset(premier_league_results_1617, select = -c(LBH,LBD,LBA,PSCH,PSCD,PSCA))

premier_league_results <- full_join(premier_league_results_1718, premier_league_results_1819, by = c("Div","Date","HomeTeam","AwayTeam","FTHG","FTAG","FTR","HTHG","HTAG","HTR","Referee","HS","AS","HST","AST","HF","AF","HC","AC","HY","AY","HR","AR","B365H","B365D","B365A","BWH","BWD","BWA","IWH","IWD","IWA","PSH","PSD","PSA","WHH","WHD","WHA","VCH","VCD","VCA","Bb1X2","BbMxH","BbAvH","BbMxD","BbAvD","BbMxA","BbAvA","BbOU","BbMx>2.5","BbAv>2.5","BbMx<2.5","BbAv<2.5","BbAH","BbAHh","BbMxAHH","BbAvAHH","BbMxAHA","BbAvAHA"))

premier_league_results <- full_join(premier_league_results, premier_league_results_1617, by = c("Div","Date","HomeTeam","AwayTeam","FTHG","FTAG","FTR","HTHG","HTAG","HTR","Referee","HS","AS","HST","AST","HF","AF","HC","AC","HY","AY","HR","AR","B365H","B365D","B365A","BWH","BWD","BWA","IWH","IWD","IWA","PSH","PSD","PSA","WHH","WHD","WHA","VCH","VCD","VCA","Bb1X2","BbMxH","BbAvH","BbMxD","BbAvD","BbMxA","BbAvA","BbOU","BbMx>2.5","BbAv>2.5","BbMx<2.5","BbAv<2.5","BbAH","BbAHh","BbMxAHH","BbAvAHH","BbMxAHA","BbAvAHA"))
```

The next step is to check if the same teams are present in all seasons.

```{r, echo=FALSE}
# Checking if same teams are present in both seasons
uniqueteams <- cbind(unique(premier_league_results_1617$HomeTeam), unique(premier_league_results_1718$HomeTeam), unique(premier_league_results_1819$HomeTeam))

uniqueteams <- uniqueteams[order(uniqueteams[,1], decreasing=FALSE),]

uniqueteams %>% knitr::kable(caption = "**Table 1: Premier League Teams**", col.names = c("Teams 2016-17", "Teams 2017-18", "Teams 2018-19")) %>% kable_styling(bootstrap_options = c("striped","hover","condensed"), fixed_thead = T) %>% kable_paper() %>%  scroll_box(width = "100%", height = "320px")

# Removing the unused datasets
rm(premier_league_results_1718, premier_league_results_1819, premier_league_results_1617)
```

We see that the teams Swansea, Stoke and West Brom are not playing in the 2018-19 season. Additionally, the teams Fulham, Wolves and Cardiff are not playing in the 2017-18 neither. We should take them out. In the remaining teams, Brighton, Huddersfield, Newcastle were not playing in the 2016-17 season. Additionally, the teams Hull, Middlesbrough and Sunderland were not playing in the following seasons neither. We should take them out.

```{r}
teams_to_remove = c('Swansea', 'Stoke', 'West Brom', 'Fulham', 'Wolves', 'Cardiff', 'Brighton', 'Huddersfield', 'Newcastle', 'Hull', 'Middlesbrough', 'Sunderland')

premier_league_results <- subset(premier_league_results, HomeTeam!='Swansea' & HomeTeam!='Stoke' & HomeTeam!='West Brom' & HomeTeam!='Fulham' & HomeTeam!='Wolves' & HomeTeam!='Cardiff' & HomeTeam!='Brighton' & HomeTeam!='Huddersfield' & HomeTeam!='Newcastle' & HomeTeam!='Hull' & HomeTeam!='Middlesbrough' & HomeTeam!='Sunderland')
      
premier_league_results <- subset(premier_league_results, AwayTeam!='Swansea' & AwayTeam!='Stoke' & AwayTeam!='West Brom' & AwayTeam!='Fulham' & AwayTeam!='Wolves' & AwayTeam!='Cardiff' & AwayTeam!='Brighton' & AwayTeam!='Huddersfield' & AwayTeam!='Newcastle' & AwayTeam!='Hull' & AwayTeam!='Middlesbrough' & AwayTeam!='Sunderland')                           

```

Now we rename the columns to make them more easily understandable.

```{r, echo=FALSE}
colnames(premier_league_results) <- c("league_division", "match_date", "home_team", "away_team", "full_time_home_goals", "full_time_away_goals", "full_time_results", "half_time_home_goals", "half_time_away_goals", "half_time_results", "referee", "home_shots", "away_shots", "home_shots_ontarget", "away_shots_ontarget", "home_fouls", "away_fouls", "home_corners", "away_corners", "home_yellow_cards", "away_yellow_cards", "home_red_cards", "away_red_cards", "bet365_home_win_odds", "bet365_draw_odds", "bet365_away_win_odds", "bet&win_home_win_odds", "bet&win_draw_odds", "bet&win_away_win_odds", "interwetten_home_win_odds", "interwetten_draw_odds", "interwetten_away_win_odds", "pinnacle_home_win_odds", "pinnacle_draw_odds", "pinnacle_away_win_odds", "williamhill_home_win_odds", "williamhill_draw_odds", "williamhill_away_win_odds", "vcbet_home_win_odds", "vcbet_draw_odds", "vcbet_away_win_odds", "no_betbrain_bookmarkers_odds", "betbrain_max_home_win_odds", "betbrain_avg_home_win_odds", "betbrain_max_draw_odds", "betbrain_avg_draw_win_odds", "betbrain_max_away_win_odds", "betbrain_avg_away_win_odds", "no_betbrain_bookmarkers_goals", "betbrain_max_over_2.5goals", "betbrain_avg_over_2.5goals", "betbrain_max_under_2.5goals", "betbrain_avg_under_2.5goals", "no_betbrain_bookmarkers_asianhandicap", "betbrain_size_handicap", "betbrain_max_home_asianhandicap_odds", "betbrain_avg_home_asianhandicap_odds", "betbrain_max_away_asianhandicap_odds", "betbrain_avg_away_asianhandicap_odds")

premier_league_results %>% kable(caption = "**Table 4: Premier League results (seasons 2016-17, 2017-18 and 2018-2019)**") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                fixed_thead = T) %>% kable_paper() %>%  scroll_box(width = "100%", height = "320px")
```

For our analysis, we will also take out variables league_division, referee and all the variables from the betting websites. These either have the same values for all observations, don't necessarily contribute to the analysis or use past information on the variable we want to predict (match result) and therefore can bias our outcome. Finally, we made sure that our final dataset doesn't contain missing values.

```{r}
premier_league_results <- premier_league_results %>% 
  select(-league_division)

premier_league_results <- premier_league_results %>% 
  select(-referee)

premier_league_results <- premier_league_results[,1:21]
```

```{r include=FALSE}
any(is.na(premier_league_results))
```

Now that we have a clean dataset, let's check out our data closer with a summary.

```{r, results = 'asis'}
dfSummary(premier_league_results, style = 'grid', plain.ascii = FALSE, tmp.img.dir = "/tmp")
```

```{r}
# Write CSV files
setwd(here::here("Data_New/"))
write_csv(premier_league_results,"premier_league_results_clean.csv")
```

