# Variable of Importance

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#  Load the cleaned dataset from the data.Rmd
setwd(here::here("Data_New/"))
premier_league_results <- read_csv("premier_league_results_clean.csv")
```

Variable importance is a method that provides a measure of the importance of each feature for the model prediction quality. It helps to analyze the link between the features and the response, through the model predictions. 

#As the teacher with DALEX Package
```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Take the NN machine model 
nnetFit

# Take the CART machine model 
classtreefit.pruned

# Take the NB machine model 
naivebayesfit

#Creating an explain object
x_train <- select(train_data1, -full_time_results)
y_train <- pull(train_data1, full_time_results)

explainer_nn <- DALEX::explain(model = nnetFit,
                                is_multiclass = TRUE,
                                data = x_train,
                                y = y_train,
                                label = "NN machine model")
  #NN fonctionne

explainer_cart <- DALEX::explain(model = classtreefit.pruned,
                                is_multiclass = TRUE,
                                data = x_train,
                                y = y_train,
                                label = "CART machine model ")
  #CART fonctionne

explainer_nb <- DALEX::explain(model = naivebayesfit,
                                is_multiclass = TRUE,
                                data = x_train,
                                y = as.numeric(y_train),
                                label = "NB machine model")
  #Problem : erreur Model info detected regression task but 'y' is a factor . (  WARNING  ), so i turn into as.numeric
  # Another error : the predict_function returns an error when executed (  WARNING  ) 
```


# Plotting the feature importance
```{r, echo = FALSE, message = FALSE, warning = FALSE}
calculate_importance <- function(your_model_explainer, n_permutations = 10) {
  imp <- model_parts(explainer = your_model_explainer,
                     B = n_permutations,
                     type = "ratio",
                     N = NULL)
  return(imp)
}

importance_nn <- calculate_importance(explainer_nn)
importance_cart <- calculate_importance(explainer_cart)
#importance_nb <- calculate_importance(explainer_nb)

plot(importance_nn, importance_cart) +
 ggtitle("Mean variable-importance ratio over 10 permutations", "")

```

# interpretation
Once we’ve identified influential variables across all three models, next we likely want to understand how the relationship between these influential variables and the predicted response differ between the models.
```{r, echo = FALSE, message = FALSE, warning = FALSE}
# compute PDP for a given variable --> uses the pdp package

```
